Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-02-20 ms.tex
working2

If the SMBH dominates the gravitational potential at the stagnation radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius resides well inside the sphere of influence ($r_{\rm s} < r_{\rm soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation (\ref{eq:stag_analytic}) simplifies to
\begin{align}
  &\rs=\frac{7}{2}\frac{G \Mbh}{n \tilde{v}_{w}^{2}} \approx 6
 \, \pc \,\, \Mbheight v_{500}^{-2}\,\,\,\,(v_{\rm w} \gtrsim \sigma),
  \label{eq:stag_simple}
\end{align}
where for the numerical estimate we adopt a typical value of $n = 1$.  This expression is similar to that obtained by \citet{Volonteri+11} on more heuristic grounds (their eq.~6).

@@

The accretion rate onto the SMBH is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc|_{r_{\rm s}}}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma$ by adopting equation (\ref{eq:stag_simple}) for $r_s$, and we have divided our results into core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies.  The resulting Eddington ratio is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.6M_{\bullet,8}M_{\odot}$ yr$^{-1}$  is the Eddington accretion rate, assuming a radiative efficiency of ten per cent.  Note the sensitive inverse dependence of the accretion rate on the wind heating rate.  
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w} \gtrsim 150M_{\bullet,8}^{0.5}$

@@

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of radius for our fiducial solutions shown in Figure \ref{fig:profiles}.  For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$ we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| < 1$ and $\tcool/\tff \lesssim 1,$ depending on the wind mass loss parameter $\eta$.  At the stagnation radius, the ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ reaches a value that lies within a factor of two of its minimum across the entire grid;  thus, the value of $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ provides a diagnostic of thermal instabilty for the entire solution.

>>>>

If the SMBH dominates the gravitational potential at the stagnation radius, then $v_{w} \gtrsim \sigma$ and hence the stagnation radius resides well inside the sphere of influence ($r_{\rm s} < r_{\rm soi}$; eq.~[\ref{eq:rsoi}]).  In this limit, equation (\ref{eq:stag_analytic}) simplifies to
\begin{eqnarray}
  &\rs \underset{v_w \gtrsim \sigma}\approx \frac{7}{2}\frac{G \Mbh}{n \tilde{v}_{w}^{2}} \approx \begin{cases}
    12
 \, \pc \,\, \Mbheight v_{500}^{-2}\, \pyear& \text{core} \\
    6
 \, \pc \,\, \Mbheight v_{500}^{-2} \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:stag_simple}
\end{eqnarray}
where for the numerical estimate we adopt values of $n \approx 1$ and $n \approx 0.5$ for core ($\Gamma = 0.1$) and cusp ($\Gamma = 0.8$) galaxies, respectively, so chosen to match the results of our numerical calculations (Fig.~\ref{fig:stag}).  This expression is similar to that obtained by \citet{Volonteri+11} on more heuristic grounds (their eq.~6).  {\bf BDM: all analytic expressions have to be updated in core case to reflect this larger rs}

@@

The accretion rate onto the SMBH is given by the integrated mass loss rate interior to the stagnation radius (eq.~[\ref{eq:q}]), 
\begin{eqnarray}
  \dot{M} &=& \frac{\eta \Menc|_{r_{\rm s}}}{\th} \nonumber \\
&\approx&
  \begin{cases}
    2.9 \times 10^{-5} M_{\bullet,8}^{1.76}
    v_{500}^{-3.8}  \eta_{0.02} \Msun \, \pyear& \text{core} \\
    5.3 \times 10^{-5} M_{\bullet,8}^{1.48} 
    v_{500}^{-2.4}  \eta_{0.02} \Msun \, \pyear  & \text{cusp}, 
  \end{cases}
  \label{eq:mdot_analytic}
\end{eqnarray}
where we have assumed $v_{w} \gg \sigma$ by adopting equation (\ref{eq:stag_simple}) for $r_s$.  The resulting Eddington ratio is given by 
\begin{eqnarray}
\frac{\dot{M}}{\dot{M}_{\rm edd}} &\approx&
  \begin{cases}
    1.3 \times 10^{-5} M_{\bullet,8}^{0.76}
    v_{500}^{-3.8}  \eta_{0.02}   & \text{core}, \\
    2.4 \times 10^{-5} \Mbheight^{0.48} 
    v_{500}^{-2.4}  \eta_{0.02}   & \text{cusp}, 
  \end{cases}
  \label{eq:eddr_analytic}
\end{eqnarray}
where $\dot{M}_{\rm edd} = 2.6M_{\bullet,8}M_{\odot}$ yr$^{-1}$  is the Eddington accretion rate, assuming a radiative efficiency of ten per cent.  Note the sensitive inverse dependence of the accretion rate on the wind heating rate.  
%To reproduce the estimated values $\dot{M}/\dot{M}_{\rm edd} \lesssim 10^{-3}$ in low-luminosity AGN or quiescent galaxies, we thus require $v_{w} \gtrsim 150M_{\bullet,8}^{0.5}$

@@

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ (eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of radius for our fiducial solutions shown in Figure \ref{fig:profiles}.  For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$ we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ and $\tcool/\tff$ can be less than unity, depending on the wind mass loss parameter $\eta$.  At the stagnation radius, the ratio $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ reaches a value that lies within a factor of two of its minimum across the entire grid;  thus, the value of $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ provides a diagnostic of thermal instabilty for the entire solution.
